### 2024-02-13 paper/SSRP/reproduction_paper.qmd
Finished SSRP reproduction paper and updated replication script

---
title: "Reproduction of \"Who Does Voter ID Keep from Voting?\" by Bernard L. Fraga and Michael G. Miller (2022)"
author: Emily Su
thanks: "Code and some of the data are available at: https://github.com/moonsdust/voting. The files for the paper that could not be uploaded to GitHub can be found here: https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/X5ALUA"
date: today
date-format: "D MMMM YYYY"
format:
  pdf:
    toc: true
number-sections: true
bibliography: references.bib
---
```{r}
#| include: false
#| warning: false
#| message: false

#### Workspace setup ####
library(tidyverse)
library(ggplot2)
library(knitr)

#### Load data ####
cleaned_reason_data <- read_csv("../../data/analysis_data/cleaned_reason_data.csv", show_col_types = FALSE)
```
# Introduction
This report contains the figures replicated from Bernard L. Fraga and Michael G. Miller's paper, "Who Does Voter ID Keep from Voting?" [@citeVoterIDPaper]. In our replication, the statistical programming language, R [@citeR], `dataverse` [@citedataverse], `tidyverse` [@citetidyverse], `ggplot2` [@citeggplot2], `haven` [@citehaven], `readr` [@citereadr], and `knitr` [@citeknitr] were used to retrieve, clean, and analyze the datasets to create the figures and tables.

# Figure A1 
The following is the replicated graph ([@fig-a1]) and a corresponding table ([@tbl-a1]) from the paper that was put online in their reproduction package [@citeVoterIDPaper]. The original graph can be found here: https://dataverse.harvard.edu/file.xhtml?fileId=4328158&version=1.0.

```{r}
#| echo: false # Code will run and its output will show but the code itself will not be included in the doc
#| warning: false # Warnings are not included 
#| message: false # Messages are not included

## Figure A1: 
# https://dataverse.harvard.edu/file.xhtml?fileId=4328158&version=1.0

figure_A1_rep_all_data <- cleaned_reason_data |>
  select(family, applied, transport, birthcert, disability, work, lost, other, black, latino, white)

# White 
figure_A1_rep_white_data <- figure_A1_rep_all_data |>
  # Filter for white
  filter(white == 1) |>
  # Remove other races
  select(-c(black, latino)) |>
  # Get percentages 
  mutate( 
    "Family Obligations" = round(sum(family) / n(), 2),
    "Applied, Has Not Recieved" = round(sum(applied) / n(), 2),
    "Lack of Transportation" = round(sum(transport) / n(), 2),
    "Lacks Necessary Documents" = round(sum(birthcert) / n(), 2), 
    "Disability/Illness" = round(sum(disability) / n(), 2),
    "Work Obligations" = round(sum(work) / n(), 2), 
    "ID was Lost or Stolen" = round(sum(lost) / n(), 2), 
    "Other Reason" = round(sum(other) / n(), 2)
  ) |>
  # Remove the original columns
  select(-c(family, applied, transport, birthcert, disability, work, lost, other, white)) |>
  # Have the percentages only show up in 1 row
  unique() |>
  # Code snippet to convert to a 1-column matrix and then to a vector 
  # was adapted from the following: https://www.geeksforgeeks.org/convert-matrix-to-vector-in-r/
  t() |> # Convert 1-row matrix into a 1-column matrix
  c() |> # Convert to a vector
  tibble() |> # Change into a tibble
  rename( # Left side is the new name and the right side is the old name
    percent = "c(...)"
  ) |>
  mutate(
    reasons = c("Family Obligations", 
                "Applied, Has Not Recieved", 
                "Lack of Transportation", 
                "Lacks Necessary Documents",
                "Disability/Illness", 
                "Work Obligations", 
                "ID was Lost or Stolen",
                "Other Reason"),
     # Add column for race
    race = "white"
  )

# Black
figure_A1_rep_black_data <- figure_A1_rep_all_data |>
  # Filter for black
  filter(black == 1) |>
  # Remove other races
  select(-c(white, latino)) |>
  # Get percentages 
  mutate( 
    "Family Obligations" = round(sum(family) / n(), 2),
    "Applied, Has Not Recieved" = round(sum(applied) / n(), 2),
    "Lack of Transportation" = round(sum(transport) / n(), 2),
    "Lacks Necessary Documents" = round(sum(birthcert) / n(), 2), 
    "Disability/Illness" = round(sum(disability) / n(), 2),
    "Work Obligations" = round(sum(work) / n(), 2), 
    "ID was Lost or Stolen" = round(sum(lost) / n(), 2), 
    "Other Reason" = round(sum(other) / n(), 2)
  ) |>
  # Remove the original columns
  select(-c(family, applied, transport, birthcert, disability, work, lost, other, black)) |>
  # Have the percentages only show up in 1 row
  unique() |>
  # Code snippet to convert to a 1-column matrix and then to a vector 
  # was adapted from the following: https://www.geeksforgeeks.org/convert-matrix-to-vector-in-r/
  t() |> # Convert 1-row matrix into a 1-column matrix
  c() |> # Convert to a vector
  tibble() |> # Change into a tibble
  rename( # Left side is the new name and the right side is the old name
    percent = "c(...)"
  ) |>
  mutate(
    reasons = c("Family Obligations", 
                "Applied, Has Not Recieved", 
                "Lack of Transportation", 
                "Lacks Necessary Documents",
                "Disability/Illness", 
                "Work Obligations", 
                "ID was Lost or Stolen",
                "Other Reason"),
    # Add column for race
    race = "black"
  )

# Latinx
figure_A1_rep_latinx_data <- figure_A1_rep_all_data |>
  # Filter for Latino
  filter(latino == 1) |>
  # Remove other races
  select(-c(white, black)) |>
  # Get percentages 
  mutate( 
    "Family Obligations" = round(sum(family) / n(), 2),
    "Applied, Has Not Recieved" = round(sum(applied) / n(), 2),
    "Lack of Transportation" = round(sum(transport) / n(), 2),
    "Lacks Necessary Documents" = round(sum(birthcert) / n(), 2), 
    "Disability/Illness" = round(sum(disability) / n(), 2),
    "Work Obligations" = round(sum(work) / n(), 2), 
    "ID was Lost or Stolen" = round(sum(lost) / n(), 2), 
    "Other Reason" = round(sum(other) / n(), 2)
  ) |>
  # Remove the original columns
  select(-c(family, applied, transport, birthcert, disability, work, lost, other, latino)) |>
  # Have the percentages only show up in 1 row
  unique() |> # Code snippet to convert to a 1-column matrix and then to a vector
  # Code snippet to convert to a 1-column matrix and then to a vector 
  # was adapted from the following: https://www.geeksforgeeks.org/convert-matrix-to-vector-in-r/
  t() |> # Convert 1-row matrix into a 1-column matrix
  c() |> # Convert to a vector
  tibble() |> # Change into a tibble
  rename( # Left side is the new name and the right side is the old name
    percent = "c(...)"
  ) |>
  mutate(
    reasons = c("Family Obligations", 
                "Applied, Has Not Recieved", 
                "Lack of Transportation", 
                "Lacks Necessary Documents",
                "Disability/Illness", 
                "Work Obligations", 
                "ID was Lost or Stolen",
                "Other Reason"),
    # Add column for race
    race = "latinx"
  )

# Combine all
# Referenced code from here: https://help.qresearchsoftware.com/hc/en-us/articles/4446615637775-How-to-Merge-Tables-Using-R
combined_figure_A1_data <- rbind(figure_A1_rep_white_data, figure_A1_rep_black_data)
combined_figure_A1_data <- rbind(combined_figure_A1_data, figure_A1_rep_latinx_data)
```

```{r}
#| label: fig-a1
#| fig-cap: Impediments of Texas voters who filed RIDs in 2016 by Race (Figure A1 from paper)
#| echo: false # Code will run and its output will show but the code itself will not be included in the doc
#| warning: false # Warnings are not included 
#| message: false # Messages are not included

# Graph
# Referenced code from https://r-graph-gallery.com/267-reorder-a-variable-in-ggplot2.html
# To figure out how to sort 
combined_figure_A1_data |>
  mutate(reasons = fct_reorder(reasons, desc(percent))) |>
  ggplot(mapping = aes(x = reasons, y = percent)) +
  geom_bar(stat="identity") +
  theme_minimal() +
  labs(
    x = "Reasons",
    y = "Percent"
  ) +
  coord_flip() + 
  # Referenced code from https://github.com/tidyverse/ggplot2/issues/2762
  # to figure out how to put labels on the left and use facet_grid
  facet_grid(vars(race), switch = "y") +
  guides(x = guide_axis(angle = 0), ) +
  theme(plot.title = element_text(size = 8, face = "bold"),
        axis.title=element_text(size = 9, face = "bold"),
        axis.text.y = element_text(size = 8)) + 
  ggtitle("Impediments of Texas voters who filed RIDs in 2016 by Race")
```
\newpage
```{r}
#| label: tbl-a1
#| tbl-cap: Impediments of Texas voters who filed RIDs in 2016 by Race (Corresponds to Figure A1 from paper)
#| echo: false # Code will run and its output will show but the code itself will not be included in the doc
#| warning: false # Warnings are not included 
#| message: false # Messages are not included 
# Table
combined_figure_A1_data |>
  kable()
```

# Figure 2
The following graph ([@fig-second]) corresponds to figure 2 in the paper's main body [@citeVoterIDPaper]. A table ([@tbl-second]) was also created to show the percentages from the replication. 
```{r}
#| echo: false # Code will run and its output will show but the code itself will not be included in the doc
#| warning: false # Warnings are not included 
#| message: false # Messages are not included

figure_2_rep_data <- cleaned_reason_data |>
  select(family, applied, transport, birthcert, disability, work, lost, other) |>
  # Get percentages 
  mutate( 
    "Family Obligations" = round((sum(family) / n()) * 100, 2),
    "Applied, Has Not Recieved" = round((sum(applied) / n()) * 100, 2),
    "Lack of Transportation" = round((sum(transport) / n()) * 100, 2),
    "Lacks Necessary Documents" = round((sum(birthcert) / n()) * 100, 2), 
    "Disability/Illness" = round((sum(disability) / n()) * 100, 2),
    "Work Obligations" = round((sum(work) / n()) * 100, 2), 
    "ID was Lost or Stolen" = round((sum(lost) / n()) * 100, 2), 
    "Other Reason" = round((sum(other) / n()) * 100, 2)
  ) |>
  # Remove the original columns
  select(-c(family, applied, transport, birthcert, disability, work, lost, other)) |>
  # Have the percentages only show up in 1 row
  unique() |>
  # Code snippet to convert to a 1-column matrix and then to a vector 
  # was adapted from the following: https://www.geeksforgeeks.org/convert-matrix-to-vector-in-r/
  t() |> # Convert 1-row matrix into a 1-column matrix
  c() |> # Convert to a vector
  tibble() |> # Change into a tibble
  rename( # Left side is the new name and the right side is the old name
    percent = "c(...)"
  ) |>
  mutate(
    reasons = c("Family Obligations", 
                "Applied, Has Not Recieved", 
                "Lack of Transportation", 
                "Lacks Necessary Documents",
                "Disability/Illness", 
                "Work Obligations", 
                "ID was Lost or Stolen",
                "Other Reason"
                )
  ) 
```

```{r}
#| label: fig-second
#| fig-cap: Impediments of Texas voters who filed RIDs in 2016 (Figure 2 from paper)
#| echo: false # Code will run and its output will show but the code itself will not be included in the doc
#| warning: false # Warnings are not included 
#| message: false # Messages are not included

# Referenced code from https://r-graph-gallery.com/267-reorder-a-variable-in-ggplot2.html
# To figure out how to sort 
# Graph
figure_2_rep_data |>
  mutate(reasons = fct_reorder(reasons, desc(percent))) |>
  ggplot(mapping = aes(x = reasons, y = percent)) +
  geom_bar(stat="identity") +
  ggtitle("Impediments of Texas voters who filed RIDS in 2016") +
  coord_flip() + 
  theme_minimal() +
  labs(x = "Reasons", y = "Percent") +
  theme(plot.title = element_text(size = 9, face = "bold"),
        axis.title = element_text(size = 9, face = "bold"))

```
\newpage
```{r}
#| label: tbl-second
#| tbl-cap: Impediments of Texas voters who filed RIDs in 2016 (Figure 2 from paper)
#| echo: false # Code will run and its output will show but the code itself will not be included in the doc
#| warning: false # Warnings are not included 
#| message: false # Messages are not included 
# Table
figure_2_rep_data |> 
  kable()
```

# Figure A3
The following is the replicated graph ([@fig-a3]) and a corresponding table ([@tbl-a3]) from the paper that was put online in their reproduction package [@citeVoterIDPaper]. The original graph can be found here: https://dataverse.harvard.edu/file.xhtml?fileId=4328141&version=1.0.
```{r}
#| echo: false # Code will run and its output will show but the code itself will not be included in the doc
#| warning: false # Warnings are not included 
#| message: false # Messages are not included

figure_A3_rep_all_data <- cleaned_reason_data |>
  select(relocation, id_capable, hardship, black, latino, white)

# White 
figure_A3_rep_white_data <- figure_A3_rep_all_data |>
  # Filter for white
  filter(white == 1) |>
  # Remove other races
  select(-c(black, latino)) |>
  # Get percentages 
  mutate( 
    "Relocation" = round(sum(relocation) / n(), 2),
    "Hardship" = round(sum(hardship) / n(), 2),
    "ID-Capable" = round(sum(id_capable) / n(), 2)
  ) |>
  # Remove the original columns
  select(-c(relocation, id_capable, hardship, white)) |>
  # Have the percentages only show up in 1 row
  unique() |>
  # Code snippet to convert to a 1-column matrix and then to a vector 
  # was adapted from the following: https://www.geeksforgeeks.org/convert-matrix-to-vector-in-r/
  t() |> # Convert 1-row matrix into a 1-column matrix
  c() |> # Convert to a vector
  tibble() |> # Change into a tibble
  rename( # Left side is the new name and the right side is the old name
    percent = "c(...)"
  ) |>
  mutate(
    reasons = c("Relocation", 
                "Hardship", 
                "ID-Capable"),
    # Add column for race
    race = "white"
  )

# Black 
figure_A3_rep_black_data <- figure_A3_rep_all_data |>
  # Filter for black
  filter(black == 1) |>
  # Remove other races
  select(-c(white, latino)) |>
  # Get percentages 
  mutate( 
    "Relocation" = round(sum(relocation) / n(), 2),
    "Hardship" = round(sum(hardship) / n(), 2),
    "ID-Capable" = round(sum(id_capable) / n(), 2)
  ) |>
  # Remove the original columns
  select(-c(relocation, id_capable, hardship, black)) |>
  # Have the percentages only show up in 1 row
  unique() |>
  # Code snippet to convert to a 1-column matrix and then to a vector 
  # was adapted from the following: https://www.geeksforgeeks.org/convert-matrix-to-vector-in-r/
  t() |> # Convert 1-row matrix into a 1-column matrix
  c() |> # Convert to a vector
  tibble() |> # Change into a tibble
  rename( # Left side is the new name and the right side is the old name
    percent = "c(...)"
  ) |>
  mutate(
    reasons = c("Relocation", 
                "Hardship", 
                "ID-Capable"),
    # Add column for race
    race = "black"
  )

# Black 
figure_A3_rep_latinx_data <- figure_A3_rep_all_data |>
  # Filter for latino
  filter(latino == 1) |>
  # Remove other races
  select(-c(white, black)) |>
  # Get percentages 
  mutate( 
    "Relocation" = round(sum(relocation) / n(), 2),
    "Hardship" = round(sum(hardship) / n(), 2),
    "ID-Capable" = round(sum(id_capable) / n(), 2)
  ) |>
  # Remove the original columns
  select(-c(relocation, id_capable, hardship, latino)) |>
  # Have the percentages only show up in 1 row
  unique() |>
  # Code snippet to convert to a 1-column matrix and then to a vector 
  # was adapted from the following: https://www.geeksforgeeks.org/convert-matrix-to-vector-in-r/
  t() |> # Convert 1-row matrix into a 1-column matrix
  c() |> # Convert to a vector
  tibble() |> # Change into a tibble
  rename( # Left side is the new name and the right side is the old name
    percent = "c(...)"
  ) |>
  mutate(
    reasons = c("Relocation", 
                "Hardship", 
                "ID-Capable"),
    # Add column for race
    race = "latinx"
  )

# Combine all
# Referenced code from here: https://help.qresearchsoftware.com/hc/en-us/articles/4446615637775-How-to-Merge-Tables-Using-R
combined_figure_A3_data <- rbind(figure_A3_rep_white_data, figure_A3_rep_black_data)
combined_figure_A3_data <- rbind(combined_figure_A3_data, figure_A3_rep_latinx_data)
```
\newpage

```{r}
#| label: fig-a3
#| fig-cap: General Impediments of Texas voters who filed RIDs in 2016 by Race (Figure A3 from paper)
#| echo: false # Code will run and its output will show but the code itself will not be included in the doc
#| warning: false # Warnings are not included 
#| message: false # Messages are not included 
# Graph
# Referenced code from https://r-graph-gallery.com/267-reorder-a-variable-in-ggplot2.html
# To figure out how to sort 
combined_figure_A3_data |>
  mutate(reasons = fct_reorder(reasons, desc(percent))) |>
  ggplot(mapping = aes(x = reasons, y = percent)) +
  geom_bar(stat="identity") +
  theme_minimal() +
  labs(
    x = "Reasons",
    y = "Percent"
  ) +
  coord_flip() + 
  # Referenced code from https://github.com/tidyverse/ggplot2/issues/2762
  # to figure out how to put labels on the left and use facet_grid
  facet_grid(vars(race), switch = "y") +
  guides(x = guide_axis(angle = 0), ) +
  theme(plot.title = element_text(size = 9, face = "bold"),
        axis.title=element_text(size = 9, face = "bold")) + 
  ggtitle("General Impediments of Texas voters who filed RIDs in 2016 by Race")
```

```{r}
#| label: tbl-a3
#| tbl-cap: General Impediments of Texas voters who filed RIDs in 2016 by Race (Figure A3 from paper)
#| echo: false # Code will run and its output will show but the code itself will not be included in the doc
#| warning: false # Warnings are not included 
#| message: false # Messages are not included

# Table
combined_figure_A3_data |>
  kable()
```
From our reproduction, we found that our results for figures A1, 2, and A3 were consistent with the figures from the the original paper [@citeVoterIDPaper].

# References